# **ADVERT CONSUMER ANALYSIS & MODELING**

## **Objective:**

To create a model that accurately predicts the likelihood of a blog reader clicking on an advert. 

## **Success Metric:**

Establish the specific groups of individuals more like to view web adverts on cryptography given the data provided.
Create a model with at least 90%  accuracy

## **The context:**

The data was collected from a blog the ran ads to advertise a course. The data contains daily time Spent on site, age of the user, area income, daily internet usage of the user, ad Topic Line,	the city, gender,country and the time stamp. The data source can be accessed from http://bit.ly/IPAdvertisingData

## **Experimental Design**

R has been used for this analysis. The exploratory data analysis has given graphical presentations of univariate and bivariate analysis.

## **Reading the data set**

```{r}
advert <-read.csv('http://bit.ly/IPAdvertisingData')
advert_df <- data.frame(advert)
rmarkdown::paged_table(head(advert_df,n=4))
```
```{r}
#Checking the the last 4 rows of the data set
rmarkdown::paged_table(tail(advert_df,n=4))
```
```{r}
#Checking the number of rows and columns in the data set
dim(advert_df)
#The data set has 1000 rows and 10 columns
```
```{r}
#Checking the data types of the columns
sapply(advert_df, class)
```
```{r}
#Checking column names
colnames(advert_df)
```
## **Data Cleaning**
**Checking for duplicates**
```{r}
sum(duplicated(advert_df))
#The data has no duplicated rows
```
**Checking for missing values**
```{r}
colSums(is.na(advert_df))
#There are no missing values in all columns
```
**Checking for outliers**
```{r}
boxplot(advert_df$Daily.Time.Spent.on.Site)
#The are no outliers on the 'daily time spent on site' column
```
```{r}
boxplot(advert_df$Age)
#The are no outliers on the 'age' column
```
```{r}
boxplot(advert_df$Area.Income)
#The are outliers on the 'area income' column. However, the outliers will not be dropped as they are legitimate  data give the context of the column data.
```
**Correcting column names**
```{r}
colnames(advert_df)
```
```{r}
#Renaming the 'male' column to 'gender' where 1 represents male and 0 female
colnames(advert_df)[7] <- 'Gender'
#Removing fullstops in the column names
colnames(advert_df)[1] <- 'Daily Time Spent on Site'
colnames(advert_df)[3] <- 'Area Income'
colnames(advert_df)[4] <- 'Daily Internet Usage'
colnames(advert_df)[5] <- 'Ad Topic Line'
colnames(advert_df)[10] <- 'Clicked on Ad'
```
```{r}
#Checking corrected column names
colnames(advert_df)
```
**Correcting Column values**
```{r}
#Changing 0 and 1 values in the gender and clicked on Ad columns in male and female values and yes and no values respectively.It will help in visualizing the outcomes.
advert_df$Gender <- ifelse(advert_df$Gender == 1,'male','female')
advert_df$`Clicked on Ad` <- ifelse(advert_df$`Clicked on Ad`==1,'yes','no')
```
```{r}
#reading the data to check if the corrections made have reflected
rmarkdown::paged_table(head(advert_df,n=5))
```
## **Exploratory Data Analysis**
### **Univariate Analysis**
**Installing libraries**
```{r}
library(tidyverse)
```
```{r}
mean(advert_df$`Daily Time Spent on Site`)
#On average, the respondents spent 65.0002 minutes on the site
```
```{r}
mean(advert_df$Age)
#On average, the respondents are aged 36 years
```
```{r}
mean(advert_df$`Daily Internet Usage`)
#On average there is an internet usage of 180.0001
```
```{r}
median(advert_df$`Area Income`)
#The median area income is 57012.3
```
```{r}
hist(advert_df$`Daily Internet Usage`, main = 'Distribution of Internet Usage', 
     xlab = 'Time spent in minutes',col = 'pink',breaks=5)
#Majority used between 200 and 250 megabytes of data
```
```{r}
x <-table(advert_df$Gender)
pie(x, main = 'Gender Distribution',)
#More females than men visited the blog website.
```
```{r}
#Distribution on countries that visited the website
c <-table(advert_df$Country)
s <-sort(c,descreasing=TRUE)[1:5]
as.matrix(s)
```

```{r}
ggplot(data =advert_df, aes(x = `Clicked on Ad`,fill= `Clicked on Ad`)) +
    geom_bar()
#The data set has an equal distribution of ad viewers with 50% having viewed the ads and 50% who did no. Therefore the data set is balanced
```
### **Bivariate Analysis**
**Clicks on Ad and Age**
```{r}
#A data frame with 'clicked on Ad ='Yes'
ad_df<-advert_df[advert_df$`Clicked on Ad`=='yes',]
ggplot(data=ad_df, aes(x=Age,fill=`Clicked on Ad`))+
  geom_histogram(position= 'stack',bins=5)

#A data frame with 'clicked on Ad ='no'
ad_df1<-advert_df[advert_df$`Clicked on Ad`=='no',]
ggplot(data=ad_df1, aes(x=Age,fill=`Clicked on Ad`))+
  geom_histogram(position= 'stack',bins=5)
#The age 40 and 50 were more likely to click on the Ad than those in the aged between(30-40)
```
**Clicks on Ad and Gender**
```{r}
ggplot(data=ad_df, aes(x=Gender,fill=`Clicked on Ad`))+
  geom_bar(position= 'dodge')
#More women clicked on the ads than men. 
```
**Clicks on Ad and Area income**
```{r}
ggplot(data=ad_df, aes(x=`Area Income`,fill=`Clicked on Ad`))+
  geom_histogram(position= 'stack',bins=10)

```
**Clicks on Ad and Daily Internet Usage**
```{r}
ggplot(data=ad_df, aes(x=`Daily Internet Usage`,fill=`Clicked on Ad`))+
  geom_histogram(position= 'stack',bins=10)
```
```{r}
sort(table(ad_df$Country),descreasing= FALSE)[1:5]

```
## **Modeling**
### **Data Preparation**
```{r}
#Encoding Gender and Clicked on ad columns
advert_df$Gender <- ifelse(advert_df$Gender == 'male',1,0)
advert_df$`Clicked on Ad` <- ifelse(advert_df$`Clicked on Ad`=='yes',1,0)
```
```{r}
#Dropping unnecessary columns
advert_df$`Ad Topic Line` <- NULL
advert_df$City <- NULL
advert_df$Country <- NULL
advert_df$Year <- NULL
advert_df$Timestamp <- NULL
```
```{r}
#renaming column gender back to male
colnames(advert_df)[5] <- 'Male'
```


```{r}
#Checking the dataset
rmarkdown::paged_table(head(advert_df,n=4))
```
**Splitting data**
```{r}
set.seed(5)
sample<- caret::createDataPartition(advert_df$`Clicked on Ad`, p = 0.8,
                                  list = FALSE,
                                  times = 1)
train <- advert_df[sample,]
test <- advert_df[-sample,]
advert_df$`Clicked on Ad` <-factor(advert_df$`Clicked on Ad`)
```
```{r}
#Checking dimensions
dim(train)
dim(test)
```
**Data Scalling**
```{r}
train[1:5]<-scale(train[1:5])
xtest<-scale(test[1:5])
```

**SVM Model**
```{r}
#Training the model
library(e1071)
model= svm(formula =`Clicked on Ad` ~ .,data = train,
           type = 'C-classification',kernel = 'linear')
```
```{r}
# Making predictions using the test data
prediction <-predict(model,xtest)
cm(table(test$`Clicked on Ad`,prediction))
```
**Naive Bayes Model**
```{r}
x<-train[,1:5]
y<-train$`Clicked on Ad`

# fiting naive bayes
set.seed(42)
model1 <- naiveBayes(`Clicked on Ad` ~ ., data = train)
model
```
```{r}
prediction1 <-predict(model1,xtest)
cm(table(test$`Clicked on Ad`,prediction1))
```

## **Observations**

1. 38 more women than men click on the ad

2. People from Czech Republic and France contributed the highest among those who visited the blog website, however people from Turkey,Ethiopia and Australia contributed the highest among those who clicked on the Ads.

3. On average , people spent 65 minutes on the blog site and an average of 180 megabyte on daily internet usage

4. The average age is 36 years

5. Persons falling under the average area income (57012.3) were more likely to click on ads.

6. Majority of those who clicked on the Ads were between the ages of 40-50 while majority of those who skipped Ads were between ages 30-40.

### **Recommendations**

1. Increase Ads friendly to people from Czech Republic and France since they visit the blog the most.

2. Increase female and male friendly content in the blog in increase  female consumption to increase number of Ad consumption

3. Make more targeted ads for people between the ages of 30 - 40 years

## **Conclusion**

The models has an accuracy of 98.5% at 95% confidence interval.It is statistically significant at a p value of 0.00 which is less than alpha= 0.5. However,the SVM model classifies better with only 3 misclassifications.

Reference :file:///C:/Users/Angela%20Nyutu/Documents/Ad%20modeling/ad-analysis-modeling.html
         